---
title: "Outlier influence on a simple linear model"
author: "Christos Tsolkas"
date: "October 21, 2020"
output: ioslides_presentation
---



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Application description 

  - The applications features a scatter plot with a fitted simple linear model.
  
  - At the side there is also the resulting residuals plot.

  - Right below there is a control panel for fixing the outlier position.

  - Finally at the bottom influence measures tables are shown that measure the influence the outlier has on the fitted model in comparison to normal points. 

## Example scatter plot


```{r echo=FALSE}
defaultW <- getOption("warn") 

options(warn = -1) 

suppressPackageStartupMessages(library(ggplot2))
suppressPackageStartupMessages(library(plotly))
outlier.x = 10
outlier.y = -3
npoints <- 100
set.seed(100)
x <- runif(npoints, min = -2, max = 2)
y <- x + rnorm(npoints, mean = 0, sd = 0.3)
df <- data.frame( x = c(x, outlier.x),
                  y = c(y, outlier.y),
                  Point = c(rep("Normal point", npoints), "Outlier point")
)

g1 <- ggplot(df, aes(x, y, col = Point)) +
  geom_point(size = 4, alpha = .4) +
  expand_limits(x = c(-4, 11), y = c(-4, 11)) +
  geom_smooth(method = "lm", formula = y ~ x,  se = FALSE, col = "deeppink") +
  theme_light() +
  theme(legend.position="bottom", legend.title = element_blank())

ggplotly(g1)

options(warn = defaultW)
```

## Idea for this app came from Residuals and Diagnostics video

<iframe width="560" height="315" src="https://www.youtube.com/embed/VohfwSJuG4k?start=94" frameborder="0" allow="accelerometer; autoplay; clipboard-write; encrypted-media; gyroscope; picture-in-picture" allowfullscreen></iframe>

## Give it a try and have fun

The application can be accessed at: [https://chrtsolkas.shinyapps.io/LinearModelDiagnostics/](https://chrtsolkas.shinyapps.io/LinearModelDiagnostics/)


